Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data

### 2018-10-09 paper.tex
added space in big data and fixed title page

\begin{document} 
    \title{Pipeline systems and infrastructure for the efficient
            and open processing of Big neuroimaging Data} 
    \author{Valerie Hayot-Sasson}
    \maketitle

@@

    \end{abstract} 
    \tableofcontents
    \chapter{Introduction}
        In recent years, the volume of neuroimaging data acquired has exceeded
        both the storage and computation capacity of a standard research 
        lab workstation. With the advancement of data sharing technologies, 
        this data has been made widely available, however, accessibility to 
        such data remains limited due to infrastructural demands and lack of 
        performant software. Even with powerful infrastructure, such as High
        Performance Computing (HPC) infrastructure, software configured to 
        efficiently use resources is required to make execution time manageable
        . Ensuring efficient processing of data is 
        complex and limits accessibility to researchers with little programming 
        knowledge. Existing neuroimaging workflow engines primarily rely on HPC
        resource managers for performance and are therefore not well suited for
        BigData processing. As a result, frameworks for the efficient 
        processing of Big neuroimaging Data need to be developed.

@@

        This paper aims to evaluate the efficiency of current neuroimaging 
        frameworks in comparison to BigData frameworks and to determine the 
        extent to which BigData frameworks must be adapted to be effective
        neuroimaging workflow engines. 
        In section~\ref{datasets}, we will briefly introduce the two 
        types of datasets that are giving rise to the neuroimaging big 
        data. Section~\ref{platforms} will provide a brief introduction 
        to the data-sharing and processing platforms available to 
        researchers. We will conclude our 
        introduction on the differences between commodity and HPC infrastructure
        in section~\ref{infrastructure}.

@@

            As a result of the rise in data sharing in neuroscience, 
            neuroimaging
            BigData has come to fruition. The field of neuroscience has now
            reached a bottleneck where there exists sufficient data to make
            impactful studies, but the software to process these datasets have 
            not evolved with the size of the data.

@@

        In this Chapter, we will compare and contrast the workflow composition 
        of seven neuroimaging frameworks (SPM, LONI, Nipype, PSOM, FastR, AA and
        Pydpiper), seven scientific workflow engines stemming from other fields
        (Taverna, Pegasus, Make, OpenMOLE, Nextflow, Galaxy, Toil), and three
        BigData frameworks (MapReduce, Spark, Dask). We will use two axes to
        evaluate workflow composition. Section~\ref{lang} will compare the 
        framework on the basis of programming language used. Section~\ref{ds}
        will then examine the differences in data structures between scientific
        and BigData frameworks. 
        Section~\ref{mod} will look at how easy it is to integrate existing 
        tools into the workflows.

@@

            As more experienced programmers may feel limited by a GUI, workflow
            engines have decided to provide users with the ability to describe
            their workflows in more flexible formats. Even those with GUIs for 
            workflow composition, such as LONI and OpenMOLE, allow more 
            experienced users to script their workflow. Four different 
            languages are used for describing scientific workflows: 1) Python, 
            2) MATLAB and GNU/Octave, 3) Shell Scripting languages, 4) Domain Specific Language (DSL). Workflows engines that 
            employ usage of Python include: Nipype, FastR, Pydpiper, Pegasus, Toil,
            MapReduce, Dask and Spark.
            Python is an excellent choice as a programming language for 
            scientific workflows, particularly neuroinformatics, due to its 
            widespread use within the community. Other advantages of 
            programming in Python include: a wide array of available libraries,
            particularly neuroimaging-specific ones such as Nilearn~\cite{nilearn} and Nibabel~\cite{matthew_brett_2018_1287921}
            and more general, but heavily used in neuroimaging, libraries such
            as SciPy~\cite{scipy} and Scikit Learn~\cite{pedregosa2011scikit},
             and an lower learning curve compared to other 
            programming languages. Additionally, Python is platform independent
            meaning it can be installed on any system. In terms of BigData 
            frameworks that employ Python, Dask is exclusively written in Python,
            whereas Hadoop MapReduce and Spark both have interfaces to Python
            while being written in Java and Scala, respectively.

@@

                The data structures made available by workflow engines dictate
                the maximum performance available to a workflow. \tristan{I don't understand the previous sentence, 
                in particular the 'maximum' idea. You should review that.} Scientific
                workflows are generally file-based, meaning that data, in the 
                form of files, will be 
                transferred across the network and I/O must be
                performed at each workflow activity. It is possible for some 
                data to remain in memory with scientific workflows, but these 
                are typically basic types such as numbers, strings and lists.
                More complex types, however, can be seen in workflow engines 
                such as PSOM which accept nested structures as long as the
                structures's terminal fields are either strings or cells of 
                strings. Due to workflow engines being file-based, performance
                is limited, particularly with BigData.

                In order to improve performance, BigData frameworks introduced
                data-centric data structures. MapReduce is key-value based, 
                therefore, data with the same keys would be located at the
                same nodes. As MapReduce still requires file-based I/O, the 
                amount of I/O remains similar to that of scientific workflows. 
                However, the key-value structures help to reduce the amount 
                of data transfers by promoting data-locality.

@@

            Unlike scientific workflow engines, BigData frameworks do not 
            provide default interfaces for the execution of complex tools 
            written in 
            languages different from their own. While Dask users can reference
            UDF Python functions directly in their workflow, Spark users
            must define a lambda function within an action or transformation 
            that executes their UDF or library on the data. Similarly to Spark,
            functions referenced in MapReduce must be called from within a Map
            or Reduce task. MapReduce can additionally reference Java 
            code and Spark can call on code also written in Scala, Java and R.

            For all workflow engines that do not, by default, interface with 
            command-line tools or containers, there are libraries that can be
            easily integrated to achieve that. 
            Boutiques~\cite{doi:10.1093/gigascience/giy016}, for instance, is a 
            Python library that can execute command-line or containerized tools
            defined within a JSON schema. Furthermore, the library's richness
            compete with Galaxy's XML. Python-based BigData frameworks can 
            easily call on this library from within the workflow in order to
            communicate with the tools.

    \chapter{Performance}\label{performance}
        Neuroimaging frameworks are notoriously known to be compute-intensive.
        The workflows and their respective tools needed to incorporate 
        parallelism in order to save on computing time. Maintaining good 
        performance has become especially critical now that neuroimaging
        BigData has emerged. The dichotomy remains between scientific 
        workflows and BigData workflows. While scientific workflows have made
        advances towards more performant engines, they still lack the 
        strategies BigData frameworks employ to reduce data transfers.

        Another notable aspect to consider is that both types of engines 
        (scientific and BigData) are tightly coupled to the infrastructure
        they were intended to be used on. Scientific workflow engines
        try to remain efficient in their handling of compute resources since such
        resources were previously scarce. In contrast, BigData frameworks 
        make efficient use of commodity clusters, where data transfers are 
        burdensome, particularly when faced with BigData.

        As some researchers are trying to shift towards the use of commodity 
        infrastructure while others are trying to process BigData on HPC, the 
        ideal workflow engine must run efficiently regardless of 
        infrastructure available. Otherwise, researchers are either left with
        workflows running inefficiently on resources or having to learn 
        multiple frameworks and rewrite their workflows depending on the 
        infrastructure selected.

        In this Chapter, we will compare scientific
        workflows to BigData frameworks on the basis of performance. 
        Section~\ref{fs} will look at what filesystems are supported by the 
        engines. This is important because data movement of any kind is costly,
        and may be worse depending on the filesystems available. Furthermore, 
        typical neuroinformatics pipelines manipulate files, and therefore, 
        reducing the amount of I/O is necessary.
        Section~\ref{sched} will provide an overview of the schedulers 
        available and their compatibility with the workflow engines. 
        Section~\ref{other} will discuss the checkpointing of data in the context
        of scientific workflows. This chapter will conclude with a discussion
        on the efficacy of Big Data frameworks for scientific computing, in
        Section~\ref{eval}.

@@

            Although MapReduce, Spark and Dask are all data-intensive worflows,
            Dask can be more similar \weird{to that of} scientific workflow, 
            particularly in a multiprocessing environment. It focuses on the 
            efficient 
            scheduling of tasks and relies on I/O and network transfers to
            communicate the data between nodes. With BigData workflows, such
            operations could incur significant performance penalties. 
            To mitigate this issue and remain efficient in the face of
            BigData, Spark and Dask implement data locality, lazy evaluation 
            and in-memory processing. Data locality is the scheduler's ability
            to schedule tasks nearest to data location to reduce data transfers.
            Lazy evaluation is the workflow's ability to optimize performance
            based on the pipeline's graph structure. In-memory processing, on 
            the other hand, tries to remove any I/O that occurs between tasks. 
            Platforms such as MapReduce rely heavily 
            on I/O, but the 
            I/O performed occurs on the compute node rather than on 
            shared storaged, making it generally more efficient than scientific
            workflows. Additionally, MapReduce uses the concept of 
            locality to ensure limit data transfers across the network.
            For the efficient processing of neuroimaging BigData, it will be
            necessary for scientific worflow engines to incorporate these 
            strategies.

            \subsection{Strategies}
                \subsubsection{In-memory computing}
                    Random Access Memory (RAM) is the fastest memory available
                    to workflow engines. BigData frameworks, such as Spark and 
                    Dask, leverage this feature by maintaining data in memory 
                    between tasks rather that saving them to local disk or 
                    shared memory which may incur significant overhead.

@@

                \subsubsection{Data locality}
                    Data locality, with respect to BigData frameworks, is a
                    scheduling
                    strategy in which location of the data is known by the 
                    scheduler and tasks are scheduled to the location where the
                    necessary data is located. It was one of the first main 
                    strategies implemented to make BigData processing more 
                    efficient. The reasoning behind data locality is that 
                    network transfers are expensive, particularly within 
                    Big Data infrastracture, and therefore transferring large 
                    volumes of data across the network for every task would be 
                    rather expensive.

@@

            \subsection{BigData Filesystem optimizations for HPC}

                Writting BigData to Lustre is very costly compared to local 
                node storage. Distributed filesystems found on HPC 
                infrastructure provide their own form of fault-tolerance, 
                therefore, adding a BigData filesystem such as HDFS which 
                replicates the data may burden the storage layer only to 
                provide a feature which already exists. This is especially 
                problematic as storage on HPC compute nodes may be limited.
                Moreover, while BigData frameworks typically execute on a 
                dedicated BigData cluster, they will be on a shared cluster in
                HPC, with many processed trying to access the memory. The 
                following are two optimizations made for the storing of BigData
                .

@@

        \section{BigData Scheduling}\label{sched}

            Scientific workflows typically rely on HPC schedulers for 
            distributed scheduling on clusters. These schedulers include 
            SLURM~\cite{yoo2003slurm},
            PBS~\cite{10.1007/3-540-60153-8_34}, 
            TORQUE~\cite{computing2015torque}, 
            SGE~\cite{SGE}, HTCondor~\cite{htcondor} and 
            Mesos~\cite{hindman2011mesos}. Additionally, they may provide
            their own form of scheduling for computations on a single node. 
            BigData frameworks, such as Dask,
            also provide interfaces to these HPC schedulers. However, other
            frameworks such as MapReduce and Spark are limited to Big Data 
            schedulers such as Yet Another Resource Negotiator (YARN) and Mesos.
            The reason MapReduce and Spark do not leverage HPC resource managers
            is that HPC resources managers are unaware of the locality of the 
            data and cannot perform any form of lazy evaluation.
            In addition to YARN and Mesos, Spark provides its own standalone 
            scheduler. The standalone scheduler is simple providing only a 
            basic scheduler. It allows users to use Spark out-of-the-box. Similarly
            to scientific workflows, all three BigData frameworks provide their
            own schedulers for single-node scheduling.

@@

            Big Data frameworks such as MapReduce, Spark and Dask do not have   
            any form of checkpointing functionality outside of a given          
            execution. For instance, it is possible to checkpoint an RDD, but   
            that RDD will only be accessible within the application's context   
            and will not be shared with future executions. As failures do       
            happen and individual neuroimaging tasks may take a non-negligable  
            amount of time to execute, a mechanism to checkpoint workflows      
            will need to be added to any BigData framework used for             
            neuroinformatics.                                                   
            \note{dont know how Taverna, Pegasus, Galaxy or Toil achieves this} 
            \note{openMole says it checks to see if outputs exist. similar to   
            make?}                                                              
            \note{what about SPM, FastR??}

        \section{Evaluation of BigData Frameworks for Scientific Computing}\label{eval}     
            The BigData frameworks have sparked significant interest within the 
            scientific community, leading to research on the efficiency of      
            these frameworks for scientific computing. ~\cite{souza2017spark} found  
            that Spark scales better than expected with longer tasks. This is   
            a consequence of there being less I/O occurring and the scheduler 
            not being as overloaded. 
            Additionally, ~\cite{boubela2016big} found that Spark in conjunction with  
            GPU acceleration greatly reduced the processing time of large       
            neuroimaging datasets.                                              
            As scientific, specifically neuroimaging workflows are typically    
            comprised of long tasks, it                                         
            is crucial that BigData frameworks scale well with tasks of longer  
            durations.

            To determine how different BigData frameworks compare in their 
            efficiency for the processing of scientific workflows, 
            ~\cite{mehta2017comparative} compared Spark, Dask and other Big Data 
            frameworks on
            a neuroimaging and astronomy use case. It was found that Dask, 
            and Spark performed consistently better than other frameworks and 
            had comparable 
            performance. Dask had poorer performance when data did not fit in
            memory. It was also found that Spark exhibited slower performance 
            when the individual images are larger due to the amount of transfers
            occuring between map and shuffle phases. Overall it was
            found that BigData Frameworks still required some adaptations to be
            effectively used as scientific workflow engines.

@@

            While all workflow engines capture generally the same type of 
            provenance information, they differ in how much provenance 
            information is captured and how it's done. LONI, Pegasus, Pydpiper 
            and Galaxy
            capture some provenance data, whereas engines 
            like Nextflow, Make, Toil and MapReduce capture very limited  
            provenance 
            information. OpenMOLE handles provenance through encapsulation of
            tasks using CARE to simulate runtime environments
            and BigData frameworks like Spark and Dask mainly rely on 
            lineage.

>>>>

\begin{document} 
    \title{Pipeline systems and infrastructure for the efficient
            and open processing of Big neuroimaging Data} 
    \author{Valerie Hayot-Sasson \\
    PhD Comprehensive Examination\\
    Departement of Computer Science and Software Engineering \\
    Concordia University \\
    Montreal, Quebec
    }
    \date{\today}
    \maketitle

@@

    \end{abstract} 
    \tableofcontents
    \chapter{Introduction}
        In recent years, the volume of neuroimaging data acquired has exceeded
        both the storage and computation capacity of a standard research 
        lab workstation. With the advancement of data sharing technologies, 
        this data has been made widely available, however, accessibility to 
        such data remains limited due to infrastructural demands and lack of 
        performant software. Even with powerful infrastructure, such as High
        Performance Computing (HPC) infrastructure, software configured to 
        efficiently use resources is required to make execution time manageable. 
        Ensuring efficient processing of data is 
        complex and limits accessibility to researchers with little programming 
        knowledge. Existing neuroimaging workflow engines primarily rely on HPC
        resource managers for performance and are therefore not well suited for
        Big Data processing. As a result, frameworks for the efficient 
        processing of Big neuroimaging Data need to be developed.

@@

        This paper aims to evaluate the efficiency of current neuroimaging 
        frameworks in comparison to Big Data frameworks and to determine the 
        extent to which Big Data frameworks must be adapted to be effective
        neuroimaging workflow engines. 
        In section~\ref{datasets}, we will briefly introduce the two 
        types of datasets that are giving rise to the neuroimaging big 
        data. Section~\ref{platforms} will provide a brief introduction 
        to the data-sharing and processing platforms available to 
        researchers. We will conclude our 
        introduction on the differences between commodity and HPC infrastructure
        in section~\ref{infrastructure}.

@@

            As a result of the rise in data sharing in neuroscience, 
            neuroimaging
            Big Data has come to fruition. The field of neuroscience has now
            reached a bottleneck where there exists sufficient data to make
            impactful studies, but the software to process these datasets have 
            not evolved with the size of the data.

@@

        In this Chapter, we will compare and contrast the workflow composition 
        of seven neuroimaging frameworks (SPM, LONI, Nipype, PSOM, FastR, AA and
        Pydpiper), seven scientific workflow engines stemming from other fields
        (Taverna, Pegasus, Make, OpenMOLE, Nextflow, Galaxy, Toil), and three
        Big Data frameworks (MapReduce, Spark, Dask). We will use two axes to
        evaluate workflow composition. Section~\ref{lang} will compare the 
        framework on the basis of programming language used. Section~\ref{ds}
        will then examine the differences in data structures between scientific
        and Big Data frameworks. 
        Section~\ref{mod} will look at how easy it is to integrate existing 
        tools into the workflows.

@@

            As more experienced programmers may feel limited by a GUI, workflow
            engines have decided to provide users with the ability to describe
            their workflows in more flexible formats. Even those with GUIs for 
            workflow composition, such as LONI and OpenMOLE, allow more 
            experienced users to script their workflow. Four different 
            languages are used for describing scientific workflows: 1) Python, 
            2) MATLAB and GNU/Octave, 3) Shell Scripting languages, 4) Domain Specific Language (DSL). Workflows engines that 
            employ usage of Python include: Nipype, FastR, Pydpiper, Pegasus, Toil,
            MapReduce, Dask and Spark.
            Python is an excellent choice as a programming language for 
            scientific workflows, particularly neuroinformatics, due to its 
            widespread use within the community. Other advantages of 
            programming in Python include: a wide array of available libraries,
            particularly neuroimaging-specific ones such as Nilearn~\cite{nilearn} and Nibabel~\cite{matthew_brett_2018_1287921}
            and more general, but heavily used in neuroimaging, libraries such
            as SciPy~\cite{scipy} and Scikit Learn~\cite{pedregosa2011scikit},
             and an lower learning curve compared to other 
            programming languages. Additionally, Python is platform independent
            meaning it can be installed on any system. In terms of Big Data 
            frameworks that employ Python, Dask is exclusively written in Python,
            whereas Hadoop MapReduce and Spark both have interfaces to Python
            while being written in Java and Scala, respectively.

@@

                The data structures made available by workflow engines dictate
                the maximum performance available to a workflow. \tristan{I don't understand the previous sentence, 
                in particular the 'maximum' idea. You should review that.} Scientific
                workflows are generally file-based, meaning that data, in the 
                form of files, will be 
                transferred across the network and I/O must be
                performed at each workflow activity. It is possible for some 
                data to remain in memory with scientific workflows, but these 
                are typically basic types such as numbers, strings and lists.
                More complex types, however, can be seen in workflow engines 
                such as PSOM which accept nested structures as long as the
                structures's terminal fields are either strings or cells of 
                strings. Due to workflow engines being file-based, performance
                is limited, particularly with Big Data.

                In order to improve performance, Big Data frameworks introduced
                data-centric data structures. MapReduce is key-value based, 
                therefore, data with the same keys would be located at the
                same nodes. As MapReduce still requires file-based I/O, the 
                amount of I/O remains similar to that of scientific workflows. 
                However, the key-value structures help to reduce the amount 
                of data transfers by promoting data-locality.

@@

            Unlike scientific workflow engines, Big Data frameworks do not 
            provide default interfaces for the execution of complex tools 
            written in 
            languages different from their own. While Dask users can reference
            UDF Python functions directly in their workflow, Spark users
            must define a lambda function within an action or transformation 
            that executes their UDF or library on the data. Similarly to Spark,
            functions referenced in MapReduce must be called from within a Map
            or Reduce task. MapReduce can additionally reference Java 
            code and Spark can call on code also written in Scala, Java and R.

            For all workflow engines that do not, by default, interface with 
            command-line tools or containers, there are libraries that can be
            easily integrated to achieve that. 
            Boutiques~\cite{doi:10.1093/gigascience/giy016}, for instance, is a 
            Python library that can execute command-line or containerized tools
            defined within a JSON schema. Furthermore, the library's richness
            compete with Galaxy's XML. Python-based Big Data frameworks can 
            easily call on this library from within the workflow in order to
            communicate with the tools.

    \chapter{Performance}\label{performance}
        Neuroimaging frameworks are notoriously known to be compute-intensive.
        The workflows and their respective tools needed to incorporate 
        parallelism in order to save on computing time. Maintaining good 
        performance has become especially critical now that neuroimaging
        Big Data has emerged. The dichotomy remains between scientific 
        workflows and Big Data workflows. While scientific workflows have made
        advances towards more performant engines, they still lack the 
        strategies Big Data frameworks employ to reduce data transfers.

        Another notable aspect to consider is that both types of engines 
        (scientific and Big Data) are tightly coupled to the infrastructure
        they were intended to be used on. Scientific workflow engines
        try to remain efficient in their handling of compute resources since such
        resources were previously scarce. In contrast, Big Data frameworks 
        make efficient use of commodity clusters, where data transfers are 
        burdensome, particularly when faced with Big Data.

        As some researchers are trying to shift towards the use of commodity 
        infrastructure while others are trying to process Big Data on HPC, the 
        ideal workflow engine must run efficiently regardless of 
        infrastructure available. Otherwise, researchers are either left with
        workflows running inefficiently on resources or having to learn 
        multiple frameworks and rewrite their workflows depending on the 
        infrastructure selected.

        In this Chapter, we will compare scientific
        workflows to Big Data frameworks on the basis of performance. 
        Section~\ref{fs} will look at what filesystems are supported by the 
        engines. This is important because data movement of any kind is costly,
        and may be worse depending on the filesystems available. Furthermore, 
        typical neuroinformatics pipelines manipulate files, and therefore, 
        reducing the amount of I/O is necessary.
        Section~\ref{sched} will provide an overview of the schedulers 
        available and their compatibility with the workflow engines. 
        Section~\ref{other} will discuss the checkpointing of data in the context
        of scientific workflows. This chapter will conclude with a discussion
        on the efficacy of Big Data frameworks for scientific computing, in
        Section~\ref{eval}.

@@

            Although MapReduce, Spark and Dask are all data-intensive worflows,
            Dask can be more similar \weird{to that of} scientific workflow, 
            particularly in a multiprocessing environment. It focuses on the 
            efficient 
            scheduling of tasks and relies on I/O and network transfers to
            communicate the data between nodes. With Big Data workflows, such
            operations could incur significant performance penalties. 
            To mitigate this issue and remain efficient in the face of
            Big Data, Spark and Dask implement data locality, lazy evaluation 
            and in-memory processing. Data locality is the scheduler's ability
            to schedule tasks nearest to data location to reduce data transfers.
            Lazy evaluation is the workflow's ability to optimize performance
            based on the pipeline's graph structure. In-memory processing, on 
            the other hand, tries to remove any I/O that occurs between tasks. 
            Platforms such as MapReduce rely heavily 
            on I/O, but the 
            I/O performed occurs on the compute node rather than on 
            shared storaged, making it generally more efficient than scientific
            workflows. Additionally, MapReduce uses the concept of 
            locality to ensure limit data transfers across the network.
            For the efficient processing of neuroimaging Big Data, it will be
            necessary for scientific worflow engines to incorporate these 
            strategies.

            \subsection{Strategies}
                \subsubsection{In-memory computing}
                    Random Access Memory (RAM) is the fastest memory available
                    to workflow engines. Big Data frameworks, such as Spark and 
                    Dask, leverage this feature by maintaining data in memory 
                    between tasks rather that saving them to local disk or 
                    shared memory which may incur significant overhead.

@@

                \subsubsection{Data locality}
                    Data locality, with respect to Big Data frameworks, is a
                    scheduling
                    strategy in which location of the data is known by the 
                    scheduler and tasks are scheduled to the location where the
                    necessary data is located. It was one of the first main 
                    strategies implemented to make Big Data processing more 
                    efficient. The reasoning behind data locality is that 
                    network transfers are expensive, particularly within 
                    Big Data infrastracture, and therefore transferring large 
                    volumes of data across the network for every task would be 
                    rather expensive.

@@

            \subsection{Big Data Filesystem optimizations for HPC}

                Writting Big Data to Lustre is very costly compared to local 
                node storage. Distributed filesystems found on HPC 
                infrastructure provide their own form of fault-tolerance, 
                therefore, adding a Big Data filesystem such as HDFS which 
                replicates the data may burden the storage layer only to 
                provide a feature which already exists. This is especially 
                problematic as storage on HPC compute nodes may be limited.
                Moreover, while Big Data frameworks typically execute on a 
                dedicated Big Data cluster, they will be on a shared cluster in
                HPC, with many processed trying to access the memory. The 
                following are two optimizations made for the storing of Big Data
                .

@@

        \section{Big Data Scheduling}\label{sched}

            Scientific workflows typically rely on HPC schedulers for 
            distributed scheduling on clusters. These schedulers include 
            SLURM~\cite{yoo2003slurm},
            PBS~\cite{10.1007/3-540-60153-8_34}, 
            TORQUE~\cite{computing2015torque}, 
            SGE~\cite{SGE}, HTCondor~\cite{htcondor} and 
            Mesos~\cite{hindman2011mesos}. Additionally, they may provide
            their own form of scheduling for computations on a single node. 
            Big Data frameworks, such as Dask,
            also provide interfaces to these HPC schedulers. However, other
            frameworks such as MapReduce and Spark are limited to Big Data 
            schedulers such as Yet Another Resource Negotiator (YARN) and Mesos.
            The reason MapReduce and Spark do not leverage HPC resource managers
            is that HPC resources managers are unaware of the locality of the 
            data and cannot perform any form of lazy evaluation.
            In addition to YARN and Mesos, Spark provides its own standalone 
            scheduler. The standalone scheduler is simple providing only a 
            basic scheduler. It allows users to use Spark out-of-the-box. Similarly
            to scientific workflows, all three Big Data frameworks provide their
            own schedulers for single-node scheduling.

@@

            Big Data frameworks such as MapReduce, Spark and Dask do not have   
            any form of checkpointing functionality outside of a given          
            execution. For instance, it is possible to checkpoint an RDD, but   
            that RDD will only be accessible within the application's context   
            and will not be shared with future executions. As failures do       
            happen and individual neuroimaging tasks may take a non-negligable  
            amount of time to execute, a mechanism to checkpoint workflows      
            will need to be added to any Big Data framework used for             
            neuroinformatics.                                                   
            \note{dont know how Taverna, Pegasus, Galaxy or Toil achieves this} 
            \note{openMole says it checks to see if outputs exist. similar to   
            make?}                                                              
            \note{what about SPM, FastR??}

        \section{Evaluation of Big Data Frameworks for Scientific Computing}\label{eval}     
            The Big Data frameworks have sparked significant interest within the 
            scientific community, leading to research on the efficiency of      
            these frameworks for scientific computing. ~\cite{souza2017spark} found  
            that Spark scales better than expected with longer tasks. This is   
            a consequence of there being less I/O occurring and the scheduler 
            not being as overloaded. 
            Additionally, ~\cite{boubela2016big} found that Spark in conjunction with  
            GPU acceleration greatly reduced the processing time of large       
            neuroimaging datasets.                                              
            As scientific, specifically neuroimaging workflows are typically    
            comprised of long tasks, it                                         
            is crucial that Big Data frameworks scale well with tasks of longer  
            durations.

            To determine how different Big Data frameworks compare in their 
            efficiency for the processing of scientific workflows, 
            ~\cite{mehta2017comparative} compared Spark, Dask and other Big Data 
            frameworks on
            a neuroimaging and astronomy use case. It was found that Dask, 
            and Spark performed consistently better than other frameworks and 
            had comparable 
            performance. Dask had poorer performance when data did not fit in
            memory. It was also found that Spark exhibited slower performance 
            when the individual images are larger due to the amount of transfers
            occuring between map and shuffle phases. Overall it was
            found that Big Data Frameworks still required some adaptations to be
            effectively used as scientific workflow engines.

@@

            While all workflow engines capture generally the same type of 
            provenance information, they differ in how much provenance 
            information is captured and how it's done. LONI, Pegasus, Pydpiper 
            and Galaxy
            capture some provenance data, whereas engines 
            like Nextflow, Make, Toil and MapReduce capture very limited  
            provenance 
            information. OpenMOLE handles provenance through encapsulation of
            tasks using CARE to simulate runtime environments
            and Big Data frameworks like Spark and Dask mainly rely on 
            lineage.
